---
title: "GSEA Analysis on non-integrated data using ClusterProfiler"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

<!-- Reference -->
<!-- https://alexslemonade.github.io/refinebio-examples/03-rnaseq/pathway-analysis_rnaseq_02_gsea.html#4_Gene_set_enrichment_analysis_-_RNA-seq -->


```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
            x)
  } else x
}
```

```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```


```{r, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
# Load required libraries
library(openxlsx)
library(clusterProfiler)
library(msigdbr)
library(org.Mm.eg.db)
library(tidyverse)
library(readxl)
library(rmarkdown)
library(ggplot2)
library(stringr)
```

<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.


R: `r R.Version()$version.string`,

openxlsx: `r packageVersion('openxlsx')`,

clusterProfiler: `r packageVersion('clusterProfiler')`, 

msigdbr: `r packageVersion('msigdbr')`, 

org.Mm.eg.db: `r packageVersion('org.Mm.eg.db')`, 

tidyverse: `r packageVersion('tidyverse')`,

readxl: `r packageVersion('readxl')`, 

rmarkdown: `r packageVersion('rmarkdown')`,

ggplot2: `r packageVersion('ggplot2')`,

stringr: `r packageVersion('stringr')`

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Prepare gene sets for Mus musculus
#Picking the same categories as used for DAVID analysis
set.seed(2024)
mouse_gobp_sets <- msigdbr(species = "Mus musculus", category = "C5", subcategory = "GO:BP")
mouse_kegg_sets <- msigdbr(species = "Mus musculus", category = "C2", subcategory = "CP:KEGG")
mouse_reactome_sets <- msigdbr(species = "Mus musculus", category = "C2", subcategory = "CP:REACTOME")

mouse_gene_sets <- rbind(mouse_gobp_sets, mouse_kegg_sets, mouse_reactome_sets)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Define the GSEA function for mouse
gsea_mouse <- function(df, condition){
  df_cluster <- df %>% dplyr::filter(cluster == condition)
  lfc_vector <- df_cluster$avg_log2FC
  names(lfc_vector) <- df_cluster$gene
  lfc_vector <- sort(lfc_vector, decreasing = TRUE)

  gsea_results <- GSEA(
    geneList = lfc_vector, # Ordered ranked gene list
    #minGSSize = 5, # Minimum gene set size
    #maxGSSize = 500, # Maximum gene set size
    eps = 0, # Boundary for calculating the p value
    seed = TRUE, # Set seed to make results reproducible
    pvalueCutoff = 1,
    pAdjustMethod = "BH", # Benjamini-Hochberg correction
    TERM2GENE = dplyr::select(mouse_gene_sets, gs_name, gene_symbol)
  )
  return(gsea_results@result)
}
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read the input CSV file
degs_all <- read_csv("data_clust0_positive_markers_jak2.csv")

# Define conditions (clusters) from input csv file
conditions <- unique(degs_all$cluster)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Perform GSEA for all conditions and collect results
gsea_results_all <- data.frame()
for (i in conditions) {
  gsea_all_tmp <- gsea_mouse(degs_all, i) %>% dplyr::mutate(cluster = i)
  gsea_results_all <- rbind(gsea_results_all, gsea_all_tmp)
}
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Split the results by condition
results_list <- list()
for (type in conditions) {
  type_results <- gsea_results_all %>% dplyr::filter(cluster == type) %>% dplyr::select(-cluster)
  results_list[[type]] <- type_results
}
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create a list of data frames for writing to Excel
write.xlsx(results_list, file = "GSEA_clusterprofiler_jak2.xlsx")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read the sheets into data frames
igg_results <- read_excel("GSEA_clusterprofiler_jak2.xlsx", sheet = "JAK2V617F_Hamster_IgG")
beta1_results <- read_excel("GSEA_clusterprofiler_jak2.xlsx", sheet = "JAK2V617F_HMb1_1")
```

## `r colorize('ClusterProfiler Results', 'blue')` {.tabset}

### Understanding the output

The output of the enrichment analysis is an excel file divided into subsheets based on clusters (JAK2 conditions) which has 11 headers and explanation on each of the headers can be found below:

1. ID: Identifier for the gene set. This is the identifier of the pathway or gene ontology term associated with the gene set.

2. Description: Description of the gene set. This typically provides additional details about the pathway or biological process represented by the gene set.

3. setSize: Size of the gene set, i.e., the number of genes included in the gene set.

4. enrichmentScore: The enrichment score computed by the gene set enrichment analysis. This score indicates the degree to which the gene set is overrepresented at the top or bottom of a ranked list of genes (based on fold changes or another ranking metric).  A higher enrichmentScore (closer to 1 or -1) indicates stronger enrichment, which is desirable as it suggests that the gene set is significantly overrepresented in the analyzed dataset. 

5. NES (Normalized Enrichment Score): The normalized enrichment score, which is the enrichment score normalized to account for differences in gene set sizes and correlations between gene sets and the data being analyzed. Higher NES is desired and considered more reliable whereas A NES close to 0 indicates no enrichment.

6. pvalue: The p-value associated with the enrichment score. It indicates the statistical significance of the enrichment of the gene set in the ranked list of genes. Lower p value is better.

7. p.adjust: The adjusted p-value after multiple testing correction. Adjusted using Benjamini-Hochberg (BH) procedure. Lower values are better.

8. qvalue: The q-value, which is another measure of significance adjusted for multiple testing. It is similar to the adjusted p-value. Lower values are better.

9. rank: The rank at which the gene set achieves its maximum enrichment score in the ranked list of genes. A lower rank indicates that the gene set is enriched towards the top of the ranked list of genes, which is desirable as it suggests higher significance. 

10. leading_edge: Information about the leading edge subset of genes contributing to the enrichment score. This typically includes the percentage of tags, list, and signal genes in the leading edge. Higher percentage contribution indicates a substantial contribution of these genes to the observed enrichment.

11. core_enrichment: Specific genes involved in the core enrichment of the gene set. These are often the key members of the gene set contributing to its enrichment score.

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Display the contents of the sheets using paged_table
cat('###', "JAK2V617F_Hamster_IgG", ' \n')
paged_table(igg_results)
cat(' \n \n')
```

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Display the contents of the sheets using paged_table
cat('###', "JAK2V617F_HMb1_1", ' \n')
paged_table(beta1_results)
cat(' \n \n')
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Define the path to your Excel file
excel_path <- "GSEA_clusterprofiler_jak2.xlsx"

# Read data from the two sheets
igg <- read_excel(excel_path, sheet = "JAK2V617F_Hamster_IgG")
beta1 <- read_excel(excel_path, sheet = "JAK2V617F_HMb1_1")

```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Function to create a bar plot for the top 10 pathways
plot_top_pathways <- function(gsea_results, title) {
  # Filter the top 10 pathways based on adjusted p-value
  top_gsea_results <- gsea_results %>%
    arrange(p.adjust) %>%
    head(10) 
  
    # Create a horizontal bar plot of top 10 enriched pathways
    ggplot(top_gsea_results, aes(x = reorder(Description, -p.adjust), y = -log10(p.adjust))) +
    geom_bar(stat = "identity", fill = "darkred") +
    coord_flip() +
    labs(title = title, x = "Pathway", y = "-log10(Adjusted P-value)") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6)) 
}
  
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Barplots for both conditions
plot_igg <- plot_top_pathways(igg, "Top 10 Pathways - JAK2V617F Hamster IgG")
plot_beta1 <- plot_top_pathways(beta1, "Top 10 Pathways - JAK2V617F HMb1_1")
```
 

## `r colorize('Barplots of top 10 pathways based on adj p value', 'blue')` {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE,fig.width=10,fig.height=6}
# Display barplots
cat('###', "JAK2V617F_Hamster_IgG", ' \n')
print(plot_igg)
cat(' \n \n')
cat('###', "JAK2V617F_HMb1_1", ' \n')
print(plot_beta1)
cat(' \n \n')
```


