---
title: "Trajectory Analysis for Oligo Cluster"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```


```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(monocle3)
library(SeuratWrappers)
```

<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

Monocle3 : `r packageVersion('monocle3')`,

SeuratWrappers: `r packageVersion('SeuratWrappers')`

# `r colorize('Trajectory/Pseudotime Analysis for Oligo Cluster using Monocle3', 'blue')`

<!-- # ```{r, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- # data.combined <- readRDS("Cluster0filt_data_Oligo_3_subclusters.RDS") -->
<!-- #  -->
<!-- # old_to_new_labels <- c( -->
<!-- #                           "Oligo-1" = "Mature Oligodendrocyte", -->
<!-- #                           "Immature-Oligo" = "Immature Oligodendrocyte", -->
<!-- #                           "TBD" = "Synaptic Oligodendrocyte" -->
<!-- #                        ) -->
<!-- # #Get the current cluster identities --> -->
<!-- # current_clusters <- Idents(data.combined) -->
<!-- # # Map the old cluster labels to the new labels --> -->
<!-- # new_cluster_ids <- plyr::mapvalues(current_clusters, from = names(old_to_new_labels), to = old_to_new_labels) -->
<!-- # # Set the new cluster identities -->
<!-- # data.combined <- SetIdent(data.combined, value = new_cluster_ids) -->
<!-- # data.combined$renamed_3_clusters <- Idents(data.combined) -->
<!-- # current_idents <- Idents(data.combined) -->
<!-- # new_idents <- factor(current_idents) -->
<!-- # Idents(data.combined) <- new_idents -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- # saveRDS(data.combined, file = "Cluster0filt_data_Oligo_3_subclusters.RDS") -->
<!-- # ``` -->


```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
data.combined <- readRDS("Cluster0filt_data_Oligo_3_subclusters.RDS")
my.so<-data.combined

# Project PC dimensions to whole data set
my.so <- ProjectDim(my.so, reduction = "pca")
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Create an expression matrix
expression_matrix <- my.so@assays$RNA@counts

# Get cell metadata
cell_metadata <- my.so@meta.data
if (all.equal(colnames(expression_matrix), rownames(cell_metadata))) {
  print(sprintf("Cell identifiers match"))
} else {
  print(sprintf("Cell identifier mismatch - %i cells in expression matrix, %i cells in metadata",
                ncol(expression_matrix), nrow(cell_metadata)))
  print("If the counts are equal, sort differences will throw this error")
}
```

```{r,eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# get gene annotations
gene_annotation <- data.frame(gene_short_name = rownames(my.so@assays$RNA), row.names = rownames(my.so@assays$RNA))
if (all.equal(rownames(expression_matrix), rownames(gene_annotation))) {
  print(sprintf("Gene identifiers all match"))
} else {
  print(sprintf("Gene identifier mismatch - %i genes in expression matrix, %i gene in gene annotation",
                nrow(expression_matrix), nrow(gene_annotation)))
  print("If the counts are equal, sort differences will throw this error")
}
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Seurat-derived CDS
cds <- new_cell_data_set(expression_matrix,
                            cell_metadata = cell_metadata,
                            gene_metadata = gene_annotation)
```


```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
#Transfer Seurat embeddings and cluster info to get same UMAP
reducedDim(cds, type = "PCA") <- my.so@reductions$pca@cell.embeddings 
cds@reduce_dim_aux$prop_var_expl <- my.so@reductions$pca@stdev
#plot_pc_variance_explained(cds)
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
cds@int_colData@listData$reducedDims$UMAP <- my.so@reductions$umap@cell.embeddings
plot_cells(cds)
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
cds@clusters$UMAP_so$clusters <- my.so@meta.data$renamed_3_clusters
cds <- cluster_cells(cds, reduction_method = "UMAP", partitioning_method = "seurat_clusters")
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
saveRDS(cds, file = "oligo_monocle3_cds.RDS")
```

## Seurat and Monocle3 clusters {.tabset}
```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
my.so <- readRDS("Cluster0filt_data_Oligo_3_subclusters.RDS")
cds <- readRDS("oligo_monocle3_cds.RDS")

#Plot clusters for comparsion between Seurat and Monocle3 clustering
cat('###',"Seurat Clusters",' \n')
DimPlot(my.so, reduction = "umap")+ggtitle("Seurat Cluster")
cat('\n \n')
cat('###',"Monocle3 Clusters",' \n')
plot_cells(cds, color_cells_by = "cluster", label_cell_groups=FALSE)+ggtitle("Monocle Cluster")
cat('\n \n')
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
cds <- learn_graph(cds)
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Extract the Seurat cluster information from the Monocle3 object
cds@clusters$renamed_3_clusters <- my.so@meta.data$renamed_3_clusters

# Identify the root cells that correspond to the Immature Oligodendrocyte cluster
root = "Immature Oligodendrocyte"
root_cells <- rownames(subset(colData(cds), renamed_3_clusters == root))
print(table(cds@clusters$renamed_3_clusters))
print(length(root_cells))
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Order cells in the Monocle3 object using the identified root cells
cds <- order_cells(cds, root_cells = root_cells)
```

```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
saveRDS(cds, file = "oligo_monocle3_cds.RDS")
```

## Monocle3 Pseudotime Analysis {.tabset}
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot cells with pseudotime
#cds <- readRDS("oligo_monocle3_cds.RDS")
seurat_pseudotime <- plot_cells(cds, reduction_method = "UMAP",cell_size = 0.5, color_cells_by = "pseudotime") + ggtitle("Monocle3 Pseudotime Analysis")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot cells with cell types
celltype_pseudotime <- plot_cells(
  cds, 
  reduction_method = "UMAP", 
  cell_size = 0.5, 
  color_cells_by = "renamed_3_clusters", 
  label_cell_groups = FALSE
) + theme(legend.position = "right") + ggtitle("Pseudotime Analysis with Cell Type")
```

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
#Plot pseudotime analysis results
cat('###',"Pseudotime",' \n')
print(seurat_pseudotime)
cat('\n \n')
cat('###',"Pseudotime with Cell Type",' \n')
print(celltype_pseudotime)
cat('\n \n')
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_oligo_trajectory_analysis.pdf"), width=10, height=8, useDingbats = FALSE)
print(seurat_pseudotime)
dev.off()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_oligo_trajectory_analysis_celltype.pdf"), width=10, height=8, useDingbats = FALSE)
print(celltype_pseudotime)
dev.off()
```
