---
title: "DEG Markers for Astrocytes Subclusters in 10 samples"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```


```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(ggrepel)
```

<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Useful string constant
suffix <- "Seurat"
sampleNames <- c("RD_VA_01", "RD_VA_02", "RD_VA_03", "RD_VA_04", "RD_VA_05", "RD_VA_06", 
                 "RD_VA_07", "RD_VA_08", "RD_VA_09", "RD_VA_10","RD_VA_11","RD_VA_12")
sample_Names_12 = gsub("_", "-", sampleNames)
rawDataDir_1 <- "/projectnb/czlab/A10_Rosene/Data/20230408_combinedTwo_results/"
rawDataDir_2 <- "/outs/filtered_feature_bc_matrix"
projectName <- "RD_VA_01_12"



# Seurat parameters
variable.features <- 2000
nPCs <- 50
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data.combined <- readRDS("data_Integrated_10_samples_5_celltypes.RDS")
DimPlot(data.combined, reduction = "umap", label = TRUE, repel = TRUE, raster=FALSE)+ ggtitle('All Samples')
```


# `r colorize('UMAP of Astrocyte Subcluster', 'blue')` 

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
#data.combined.subset <- subset(data.combined, subset = (seurat_5_clusters == "Astrocyte"))
data.combined <- readRDS("astrocyte_subcluster.RDS")
DimPlot(data.combined, reduction = "umap")+ ggtitle('Astrocyte Subcluster')
```


```{r, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_astrocyte_cluster_umap.pdf"), width=10, height=8, useDingbats = FALSE)
DimPlot(data.combined, reduction = "umap")+ ggtitle('Astrocyte Subcluster')
dev.off()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data.combined <- readRDS ("astrocyte_subcluster.RDS")
cat(paste0("Number of cells in Astrocyte cluster: " , ncol(data.combined)))
```

<!-- # ```{r, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- # saveRDS(data.combined.subset, file = "astrocyte_subcluster.RDS") -->
<!-- # ``` -->


# `r colorize('Identifying markers in Astrocyte Cluster', 'blue')` {.tabset}

<!-- # ```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- # data.combined <- readRDS("astrocyte_subcluster.RDS") -->
<!-- # Idents(data.combined) <- "group" -->
<!-- #  -->
<!-- #  -->
<!-- # # Find all markers for immature oligo -->
<!-- # DefaultAssay(data.combined) <- "RNA" -->
<!-- # data.markers <- FindAllMarkers(data.combined, only.pos = FALSE, min.pct = 0, logfc.threshold = 0, return.thresh=1) -->
<!-- #  -->
<!-- # # Save markers -->
<!-- # write.csv(data.markers, file='data_group_volcano_deg_markers_astrocyte.csv') -->
<!-- # ``` -->



```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Markers .csv
data.markers = read.csv('data_group_markers_astrocyte.csv')
data.markers$pct.diff = data.markers$pct.1 - data.markers$pct.2
data.markers$pct.ratio = data.markers$pct.1 / data.markers$pct.2
avglog_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = avg_log2FC)
avglog_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = avg_log2FC)
pctdiff_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.diff)
pctdiff_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.diff)
pctratio_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.ratio)
pctratio_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.ratio)
```


```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}

positive.markers = read.csv('data_group_markers_astrocyte.csv')

positive.markers$pct.diff = positive.markers$pct.1 - positive.markers$pct.2
positive.markers$pct.ratio = positive.markers$pct.1 / positive.markers$pct.2

# Keep positive markers only
positive.markers <- positive.markers[positive.markers$avg_log2FC > 0.0,]

positive.markers <- positive.markers %>%
  arrange(p_val) %>%     # Arrange by p_val in ascending order
  group_by(cluster) %>%  # Group by cluster
  slice_head(n = 200000) # Select top 200000 rows per group

write.csv(positive.markers, file='data_positive_group_markers_astrocyte.csv')

```

```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
paged_table(positive.markers)
```

# Positive markers for Astrocyte Cluster {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('##',"Top 20 Markers (avg_log2FC)",' \n')
paged_table(avglog_p)
cat(' \n \n')
cat('##',"Top 20 Markers (pct.diff)",' \n')
paged_table(pctdiff_p)
cat(' \n \n')
cat('##',"Top 20 Markers (pct.ratio)",' \n')
paged_table(pctratio_p)
cat(' \n \n')

```


# Negative markers for Astrocyte Cluster {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('##',"Top 20 Markers (avg_log2FC)",' \n')
paged_table(avglog_n)
cat(' \n \n')
cat('##',"Top 20 Markers (pct.diff)",' \n')
paged_table(pctdiff_n)
cat(' \n \n')
cat('##',"Top 20 Markers (pct.ratio)",' \n')
paged_table(pctratio_n)
cat(' \n \n')

```


# All markers for Astrocyte Cluster {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
paged_table(data.markers)
```


# `r colorize('Dot Plots of Top 30 markers for Astrocyte Subcluster', 'blue')` {.tabset}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Dot plot for astrocyte cluster

#load subset data
data.combined <- readRDS("astrocyte_subcluster.RDS")
Idents(data.combined) <- "group"

# Read positive marker data
data.markers <- read.csv("data_group_markers_astrocyte.csv")

# Group by cluster and select top 30 markers based on avg_log2FC for one condition
top_markers <- data.markers %>%
  group_by(cluster) %>%
  slice_max(n = 30, order_by = avg_log2FC) %>%
  ungroup()

# Sort the genes for each condition for the dotplot
top_markers_sorted <- top_markers %>%
  arrange(cluster, desc(avg_log2FC))

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load subset data
#data.combined <- readRDS("astrocyte_subcluster.RDS")
DefaultAssay(data.combined) <- "RNA"

# Create a list to store dot plots
dot_plots <- list()


# Loop over each condition to create dot plots
for (cond in unique(top_markers$cluster)) {
  # Extract top genes for the current condition
  genes <- top_markers_sorted$gene[top_markers_sorted$cluster == cond]

  # Create dot plot for current condition
  dot_plot <- DotPlot(object = data.combined,
                      features = genes,
                      dot.scale = 7) +
              RotatedAxis() +
              theme(axis.text.y = element_text(size = 8),
                    axis.text.x = element_text(size = 8)) +
              labs(title = paste("Astrocyte Cluster Markers",cond)) +
              scale_colour_gradient2(low = "white",mid="darkgrey", high = "darkblue")

  # Store the dot plot in the list
  dot_plots[[paste(cond)]] <- dot_plot
}

```



```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Print all dot plots
cat('##',"CR Markers",' \n')
print(dot_plots["CR"])
cat('\n \n')
cat('##',"CONTROL Markers",' \n')
print(dot_plots["CONTROL"])
cat('\n \n')
```


# `r colorize('Volcano plot for DEGs for Astrocyte Cluster', 'blue')` {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}

# Read the DEGs from the CSV files
de <- read.csv('data_group_volcano_deg_markers_astrocyte.csv')
# List of oligo types
types <- c("CR")

de<- subset(de, cluster=="CR")
de$diffexpressed <- "NO"

de$diffexpressed[de$avg_log2FC < -0.25 & de$p_val < 0.05] <- "DOWN"
de$diffexpressed[de$avg_log2FC > 0.25 & de$p_val < 0.05] <- "UP"

subset_df <- de[complete.cases(de[, "gene"]), ]

c1 <- subset_df[subset_df[, "avg_log2FC"] > 0 & subset_df[, "p_val"] <0.05, ] # condition 1
c2 <- subset_df[subset_df[, "avg_log2FC"] < 0 & subset_df[, "p_val"] <0.05, ] # condition 2
c1_byavg_log2FC <- c1[order(c1$avg_log2FC,decreasing = TRUE),]
c2_byavg_log2FC <- c2[order(c2$avg_log2FC,decreasing = FALSE),]
c1_top10 <- head(c1_byavg_log2FC, 10)$gene
c2_top10 <- head(c2_byavg_log2FC, 10)$gene
labeled_proteins <- c(c1_top10, c2_top10)
de$delabel <- ifelse(de$gene %in% labeled_proteins & de$diffexpressed %in% c("UP", "DOWN"), de$X, NA)

volcanoplot <- function() {

  ggplot(data = de, aes(x = avg_log2FC, y = -log10(p_val), col = diffexpressed, label = delabel)) +
  geom_vline(xintercept = c(-0.25, 0.25), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
    geom_point(data = de %>% filter(diffexpressed == "DOWN"), color = "#00AFBB",size=0.8, alpha=0.5) +
    geom_point(data = de %>% filter(diffexpressed == "UP"), color = "#bb0c00",size=0.8, alpha=0.5) +
    geom_point(data = de %>% filter(diffexpressed == "NO"), color = "gray",size=0.8, alpha=0.5) +
    geom_text_repel(min.segment.length = 0, size = 2, max.overlaps = Inf, show.legend = F,) +
  scale_color_manual(values = c("DOWN" = "black", "UP" =  "black")) +#, limits= c("UP in zero-Trp", "UP in high-Trp")) + # to set the colours of our variable
                     # labels = c("UP in zero-Trp", "Not significant", "UP in high-Trp")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  guides(color = guide_legend(title = NULL))+
  coord_cartesian(ylim = c(0, 20), xlim = c(-2, 2)) + # set limits
  labs(#color = 'Differential Expression', #legend_title
       x = expression("Upregulated in CONTROL <- avg log"[2]*"FC -> Upregulated in CR"), y = expression("-log"[10]*"p-value")) + 
  scale_x_continuous(breaks = seq(-10, 10, 1)) + # to customize the breaks in the x axis
  ggtitle(paste0('DEGs in Astrocyte Subcluster CR vs CONTROL')) + # Plot title 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text=element_text(size=16))
  
}

p <- volcanoplot() + theme(legend.position = "none")
print(p)
```

```{r, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_astrocyte_dot_plots.pdf"), width=10, height=8, useDingbats = FALSE)
print(dot_plots)
dev.off()
```


```{r, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_astrocyte_deg_volcano_plots.pdf"), width=10, height=8, useDingbats = FALSE)
print(p)
dev.off()
```
