---
title: 'PBMC23K_rm25pct FastMNN Integration'
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```


```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(batchelor)
library(SingleCellExperiment)
library(scater)
library(Matrix)
library(scran)
```
<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

batchelor: `r packageVersion('batchelor')`,

SingleCellExperiment: `r packageVersion('SingleCellExperiment')`,

scater: `r packageVersion('scater')`,

Matrix: `r packageVersion('Matrix')`,

scran: `r packageVersion('scran')`

```{r, echo=FALSE}
# Useful string constant
suffix <- "fastMNN"
batch_names <- c("P8_0", "P8_3", "P8_7")
rawDataDir <- "/projectnb/czlab/A00/ycliu/scRNA_integration_benchmark_datasets/PBMC23K_rm25pct"
projectName <- "fastmnn_PBMC23K_rm25pct"

# Seurat parameters
variable.features <- 2000
nPCs <- 50
```

```{r, eval = FALSE, echo=FALSE}
# Load data

counts <- readMM(paste0(rawDataDir, "/matrix.mtx.gz"))
genes <- read.table(paste0(rawDataDir, "/features.tsv.gz"), header=FALSE, stringsAsFactors=FALSE)
barcodes <- read.table(paste0(rawDataDir, "/barcodes.tsv.gz"), header=FALSE, stringsAsFactors=FALSE)

# Create SingleCellExperiment object
sce <- SingleCellExperiment(assays=list(counts=as.matrix(counts)))
rownames(sce) <- genes$V2
colnames(sce) <- barcodes$V1

```

```{r, eval=FALSE, echo=FALSE}

# Read in batch and label information
batch <- read.csv(paste0(rawDataDir, "/", "batch.csv"), header=FALSE)
labels <- read.csv(paste0(rawDataDir, "/", "labels.csv"), header=FALSE)
sce$batch <- as.factor(batch$V1)
sce$labels <- as.factor(labels$V1)

print(unique(sce$batch))
print(unique(sce$labels))

```

```{r,eval=FALSE, echo=FALSE}
# Verify that none of the cells got filtered while reading in data
cat("Number of cells in SingleCellExperiment object:", ncol(sce), "\n")
cat("Length of labels:", nrow(labels), "\n")
```

```{r,eval=FALSE, echo=FALSE}
# Save raw data
saveRDS(sce, file = "data_raw.RDS")
```

```{r, eval = FALSE, echo=FALSE}
# for reproducibility
set.seed(1234)

#Read raw data
sce <- readRDS("data_raw.RDS")
```

```{r, eval = FALSE, echo=FALSE}
# Normalise data
sce <- logNormCounts(sce)
# Feature selection
dec <- modelGeneVar(sce)
#Top 2000 features which will be used later in FastMNN, since it requires an equal number of genes in batches.
chosen.hvgs <- getTopHVGs(dec, n = 2000)
```

```{r,eval=FALSE, echo=FALSE}
# Perform PCA
sce <- runPCA(sce, ncomponents = 50)
# Perform UMAP
sce <- runUMAP(sce, dimred = "PCA", ncomponents = 2)
```

```{r,eval=FALSE, echo=FALSE}
# Save raw data with umap
saveRDS(sce, file = "data_raw_umap.RDS")
```

# `r colorize('Non-integrated data for PBMC23K_rm25pct', 'blue')`

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Load raw data
sce <- readRDS("data_raw_umap.RDS")
plotReducedDim(sce, dimred = "UMAP", colour_by = "batch") + ggtitle('3 Batches before Integration')

```


```{r, eval = set_eval, echo=FALSE}
# Load raw data for data integration
sce <- readRDS("data_raw_umap.RDS")

# Run fastMNN integration after subsetting for common genes found earlier)
sce_integrated <- fastMNN(sce, batch=sce$batch, subset.row=chosen.hvgs)

# Save integrated data
saveRDS(sce_integrated, file = "data_Integrated_fastMNN.RDS")
```

```{r, eval = set_eval, echo=FALSE}
# Load integrated data
sce_integrated <- readRDS("data_Integrated_fastMNN.RDS")
```

```{r, eval = set_eval, echo=FALSE}
# Run the standard workflow for visualization
sce_integrated <- runUMAP(sce_integrated, dimred="corrected",ncomponents = 2)

# Save integrated data with UMAP
saveRDS(sce_integrated, file = "data_Integrated_fastMNN_umap.RDS")
```


# `r colorize('UMAP for 3 batches after integration', 'blue')` 

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Load integrated data
sce_integrated <- readRDS("data_Integrated_fastMNN_umap.RDS")
#Alternative is plotUMAP, so dimred is not required
plotReducedDim(sce_integrated, dimred = "UMAP", colour_by = "batch") + ggtitle('3 Batches after Integration')

```

```{r, eval = set_eval, echo=FALSE}
sce_integrated <- readRDS("data_Integrated_fastMNN_umap.RDS")
```


```{r, eval = set_eval, echo=FALSE}
# Extract corrected PCA embeddings
pca.embeddings <- reducedDim(sce_integrated, "corrected")
pca.df <- as.data.frame(pca.embeddings)

# Save to CSV with index
write.csv(pca.df, file = "fastMNN_matrix_with_index.csv")

# Save to CSV without index
pca.df_no_colnames <- pca.df
colnames(pca.df_no_colnames) <- NULL
write.csv(pca.df_no_colnames, file = "fastMNN_matrix.csv", row.names = FALSE)


```

```{r, eval = set_eval, echo=FALSE}
# Extract corrected UMAP embeddings
umap.embeddings <- reducedDim(sce_integrated, "UMAP")
umap.df <- as.data.frame(umap.embeddings)

# Save to CSV with index
write.csv(umap.df, file = "fastMNN_umap_with_index.csv")

# Save to CSV without index
umap.df_no_colnames <- umap.df
colnames(umap.df_no_colnames) <- NULL
write.csv(umap.df_no_colnames, file = "fastMNN_umap.csv", row.names = FALSE)
```

