---
title: 'PBMC23K_rm25pct Seurat V4 Integration'
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```


```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
```
<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`.

```{r, echo=FALSE}
# Useful string constant
suffix <- "Seurat"
batch_names <- c("P8_0", "P8_3", "P8_7")
rawDataDir <- "/projectnb/czlab/A00/ycliu/scRNA_integration_benchmark_datasets/PBMC23K_rm25pct"
projectName <- "seuratv4_PBMC23K_rm25pct"

# Seurat parameters
variable.features <- 2000
nPCs <- 50
```

```{r, eval = FALSE, echo=FALSE}
#Read in raw data and create Seurat object with min.cell=0 and min.features=0 to avoid any default filtering
all.raw.data <- Read10X(data.dir = rawDataDir)
data <- CreateSeuratObject(counts = all.raw.data, project = projectName, min.cell=0, min.features=0)
print(unique(data@meta.data$orig.ident))
```
```{r, eval=FALSE, echo=FALSE}
#Read in batch and label information
batch <- read.csv(paste0(rawDataDir, "/", "batch.csv"), header=FALSE)
labels <- read.csv(paste0(rawDataDir, "/", "labels.csv"), header=FALSE)
```

```{r, eval=FALSE, echo=FALSE}

# Add batch and label metadata information to Seurat object
data$batch <- as.factor(batch$V1) 
data$labels <- as.factor(labels$V1)


print(unique(data$batch))
print(unique(data$labels))
```

```{r,eval=FALSE, echo=FALSE}
#Verify that none of the cells got filtered during CreateSeuratObject()

# Check the number of cells in the Seurat object
num_cells <- length(Cells(data))

# Check the length of labels 
num_labels <- nrow(labels)

cat("Number of cells in Seurat object:", num_cells, "\n")
cat("Length of labels:", num_labels, "\n")
```

```{r,eval=FALSE, echo=FALSE}
# Save raw data
saveRDS(data, file = paste("data_raw.RDS") )
```

```{r, eval = FALSE, echo=FALSE}
data <- NormalizeData(data, normalization.method = "LogNormalize", scale.factor = 10000)
data <- FindVariableFeatures(data, selection.method = "vst", nfeatures = 2000)
data <- ScaleData(data, verbose = FALSE)

# dimension reduction
data <- RunPCA(data, npcs = 50, verbose = FALSE)
data <- RunUMAP(data, reduction = "pca", dims = 1:50)

saveRDS(data, file = paste("data_raw_umap.RDS") )
```

# `r colorize('Non-integrated data for PBMC23K_rm25pct', 'blue')` {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Load raw data
data <- readRDS("data_raw_umap.RDS")
DimPlot(data, reduction = "umap", group.by= 'batch' ) + ggtitle('3 Batches before Integration')

```


```{r, eval = set_eval, echo=FALSE}
# Load raw data for data integration
data <- readRDS("data_raw_umap.RDS")
data$group <- data$batch

# split the dataset into a list of two seurat objects (stim and CTRL)
data.list = SplitObject(data, split.by = "batch")

# normalize and identify variable features for each dataset independently
data.list <- lapply(X = data.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = data.list)
```

```{r, eval = set_eval, echo=FALSE}
# Perform integration
# We then identify anchors using the FindIntegrationAnchors() function, which takes a list 
# of Seurat objects as input, and use these anchors to integrate the two datasets together with IntegrateData().
data.anchors <- FindIntegrationAnchors(object.list = data.list, anchor.features = features)

# this command creates an 'integrated' data assay
data.combined <- IntegrateData(anchorset = data.anchors)

# Save raw data
saveRDS(data.combined, file = paste("data_Integrated_3_batches.RDS") )
```

```{r, eval = set_eval, echo=FALSE}
# Load raw data
data.combined <- readRDS("data_Integrated_3_batches.RDS")
```

```{r, eval = set_eval, echo=FALSE}
# Run the standard workflow for visualization and clustering
data.combined <- ScaleData(data.combined, verbose = FALSE)
data.combined <- RunPCA(data.combined, npcs = 50, verbose = FALSE)
data.combined <- RunUMAP(data.combined, reduction = "pca", dims = 1:50)

# Save integrated data
saveRDS(data.combined, file = paste("data_Integrated_3_batches_umap.RDS") )
```

# `r colorize('UMAP for 3 batches after integration', 'blue')` {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Load integrated data
data.combined <- readRDS("data_Integrated_3_batches_umap.RDS")

DimPlot(data.combined, reduction = "umap", group.by= 'batch' ) + ggtitle('3 Batches after Integration')

```

```{r, eval = set_eval, echo=FALSE}
#data.combined <- readRDS("data_Integrated_3_batches_umap.RDS")
```

```{r, eval = set_eval, echo=FALSE}
# Extract PCA embeddings
pca.embeddings <- Embeddings(data.combined, reduction = "pca")
pca.df <- as.data.frame(pca.embeddings)

# Save to CSV with index
write.csv(pca.df, file = "seuratv4_matrix_with_index.csv")

# Save to CSV without index
pca.df_no_colnames <- pca.df
colnames(pca.df_no_colnames) <- NULL
write.csv(pca.df_no_colnames, file = "seuratv4_matrix.csv", row.names = FALSE)

```

```{r, eval = set_eval, echo=FALSE}
# Extract UMAP embeddings
umap.embeddings <- Embeddings(data.combined, reduction = "umap")
umap.df <- as.data.frame(umap.embeddings)

# Save to CSV with index
write.csv(umap.df, file = "seuratv4_umap_with_index.csv")

# Save to CSV without index
umap.df_no_colnames <- umap.df
colnames(umap.df_no_colnames) <- NULL
write.csv(umap.df_no_colnames, file = "seuratv4_umap.csv", row.names = FALSE)
```

