---
title: "MS_2-6 non-integrated data reanalysis"
author: "Saumya"
date: "06/07/2024"
output:
  html_document:
    df_print: paged
---


```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```


```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
```

<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.

dplyr: `r packageVersion('dplyr')` <br>
Seurat: `r packageVersion('Seurat')` <br>
patchwork: `r packageVersion('patchwork')` <br>
ggplot2: `r packageVersion('ggplot2')` <br>
cowplot: `r packageVersion('cowplot')` <br>



```{r, echo=FALSE}
# Useful string constant
suffix <- "Seurat"
sampleNames <- c("MS_2", "MS_3", "MS_4", "MS_5", "MS_6")
SAMPLE_GROUP = c('JAK2V617F_HMb1_1', 'C57BL6J_Hamster_IgG', 'C57BL6J_HMb1_1', 'JAK2V617F_Hamster_IgG', 'C57BL6J_HMb1_1')

sample_Names_6 = gsub("_", "-", sampleNames)
rawDataDir_1 <- "/projectnb/czlab/A03_Shinubo/Data/202305_JAK2_10X/"
rawDataDir_2 <- "/outs/filtered_feature_bc_matrix"
projectName <- "re-MS_1-6"

# Seurat parameters
variable.features <- 2000
nPCs <- 50
```


### `r colorize('Note: This analysis is based on the Non-integrated 5 samples before QC process. These 5 samples are divided into four groups:', 'red')` 

```{r,  echo=FALSE}
df <- data.frame(
  Sample = sampleNames,
  Group = SAMPLE_GROUP
)
print(df)
```



```{r, eval = FALSE, echo=FALSE}
# Load data
samples <- c()
for (sample in sampleNames) {
  name <- paste(rawDataDir_1, sample, rawDataDir_2, sep = "")
  samples <- c(samples, name)
}

names(samples) = sample_Names_6 ## No dashes should be in the sample name as that conflicts with Seurat's nomenclature
all.raw.data <- Read10X(data.dir = samples)
data <- CreateSeuratObject(counts = all.raw.data, project = projectName,  min.cells = 3, min.features = 200)#, min.cells = seuratQCMinCells, min.features = seuratQCMinGenes,)

data$sample <- data@meta.data$orig.ident

sample_to_group <- c("MS_2" = "JAK2V617F_HMb1_1",
                      "MS_3" = "C57BL6J_Hamster_IgG",
                      "MS_4" = "C57BL6J_HMb1_1",
                      "MS_5" = "JAK2V617F_Hamster_IgG",
                      "MS_6" = "C57BL6J_HMb1_1")

data$group <- factor(sample_to_group[data$sample])

# Save raw data
saveRDS(data, file = paste("data_raw.RDS") )
```



```{r, eval = FALSE, echo=FALSE}
data <- NormalizeData(data, normalization.method = "LogNormalize", scale.factor = 10000)
data <- FindVariableFeatures(data, selection.method = "vst", nfeatures = 2000)
data <- ScaleData(data, verbose = FALSE)

# dimension reduction
data <- RunPCA(data, npcs = 50, verbose = FALSE)
data <- RunUMAP(data, reduction = "pca", dims = 1:50)

saveRDS(data, file = paste("data_raw_umap.RDS") )
```


```{r, echo=FALSE}
# Load raw data
data <- readRDS("data_raw_umap.RDS")
```

# `r colorize('Non-integrated data for 5 samples(Before QC)', 'blue')` {.tabset}


```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('##','All samples',' \n')
DimPlot(data, reduction = "umap", group.by= 'sample' ) + ggtitle('All Samples')
cat(' \n \n')

for (samp in sample_Names_6) {
  cat('##',samp,' \n')
  print(DimPlot(subset(data, subset = (sample %in% c(samp))), reduction = "umap") + ggtitle(samp))
  cat(' \n \n')
}
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

data$nFeature_label <- factor(ifelse(data$nFeature_RNA > 7000, "Above 7,000", "Below 7,000"))

data$nRead_label <- factor(ifelse(data$nCount_RNA > 50000, "Above 50,000", "Below 50,000"))

# check Mitochondrial gene names
gene_names <- rownames(data)
mt_genes <- grep("^mt-", gene_names, value = TRUE)
print(mt_genes)


data[["percent.mt"]] <- PercentageFeatureSet(data, pattern = "^mt-")

# Add a new metadata column with labels based on the threshold of 10%
data$mt_label <- factor(ifelse(data$percent.mt > 10, "Above 10pct", "Below 10pct"))

```



# `r colorize('Mk-primed gene set (Rodriguez-Fraticelli et al., 2018a)', 'blue')` {.tabset}


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library("xlsx")
# ###Fourteen_clst_markers = read.xlsx('shinobu_2023-03-02_trial_run_analysis.xlsx', sheetIndex = 1)
MK_markers = read.xlsx('shinobu_TableS2_GSA_gene_set-2.xlsx', sheetIndex = 'Table S2A')

library('stringr')

MK_markers_selected = str_to_title(MK_markers$A)

gene_names <- rownames(data)

mask <- MK_markers_selected %in% gene_names
gene_selected = MK_markers_selected[mask]

```


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}


DefaultAssay(data) <- "RNA"
for (feature in gene_selected) {
  cat('##',feature,' \n')
  print(FeaturePlot(data, features = c(feature),raster=FALSE, slot = "data"))
  print(VlnPlot(data, features = c(feature), pt.size = 0.0,  assay = "RNA",
                group.by='nFeature_label')+ggtitle(paste(feature, '/ nFeature')))
  print(VlnPlot(data, features = c(feature), pt.size = 0.0,  assay = "RNA", 
                group.by='nRead_label')+ggtitle(paste(feature, '/ nRead')))
  cat(' \n \n')
}
```


# `r colorize('Checking the quality of cells', 'blue')` {.tabset}

## Number of features
```{r, echo=FALSE}
VlnPlot(data, features = c("nFeature_RNA"), pt.size = 0.0, group.by= 'sample') + ggtitle('Number of features')
```

## Number of reads
```{r, echo=FALSE}
VlnPlot(data, features = c("nCount_RNA"), pt.size = 0.0, group.by= 'sample') + ggtitle('Number of reads')
```

## Percantage of mitochondrial genes
```{r, echo=FALSE}
VlnPlot(data, features = c("percent.mt"), pt.size = 0.0, group.by= 'sample') + ggtitle('Percentage of mitochondrial genes')
```

# `r colorize('QC (Mitochondrial genes percentage)', 'blue')` {.tabset}

We find a subset of cells (the red part in UMAP) that expressing mitochondrial genes highly (Above 10%). I will remove those cells if you don't have any questions. We won't do QC based on the number of features or reads in each cell.

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('##','All samples',' \n')
DimPlot(data, reduction = "umap", group.by= 'mt_label' ) + ggtitle('All Samples')
cat(' \n \n')

for (samp in sample_Names_6) {
  cat('##',samp,' \n')
  print(DimPlot(subset(data, subset = (sample %in% c(samp))), group.by= 'mt_label', reduction = "umap") + ggtitle(samp))
  cat(' \n \n')
}
```

# `r colorize('QC (Ribosomal genes percentage)', 'blue')` {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# 
# # check Ribosomal gene names
gene_names <- rownames(data)
rp_genes <- grep("^Rp", gene_names, value = TRUE)
print(rp_genes)

data[["percent.rp"]] <- PercentageFeatureSet(data, pattern = "^Rp")

# Add a new metadata column with labels based on the threshold of 10%
data$rp_label <- factor(ifelse(data$percent.rp > 10, "Above 10pct", "Below 10pct"))

saveRDS(data, file = paste("data_raw_umap.RDS") )
```
## Percantage of ribosomal genes
```{r, echo=FALSE}
VlnPlot(data, features = c("percent.rp"), pt.size = 0.0, group.by= 'sample') + ggtitle('Percentage of ribosomal genes')
```