---
title: 'MS_2-6 integrated data step1: Do integration'
author: "K4"
date: "06/07/2024"
output:
  html_document:
    df_print: paged
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```


```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
```
<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`.


```{r, echo=FALSE}
# Useful string constant
suffix <- "Seurat"
sampleNames <- c("MS_2", "MS_3", "MS_4", "MS_5", "MS_6")
sample_Names_cp = gsub("_", "-", sampleNames)
rawDataDir_1 <- "/projectnb/czlab/A03_Shinubo/Data/202305_JAK2_10X/"
rawDataDir_2 <- "/outs/filtered_feature_bc_matrix"
projectName <- "MS_2-6"

# Seurat parameters
variable.features <- 2000
nPCs <- 50
```

### `r colorize('Note', 'blue')` 


```{r, eval = set_eval, echo=FALSE}
# Load raw data
data <- readRDS("data_raw_umap.RDS")
data$group <- data@meta.data$orig.ident

# split the dataset into a list of two seurat objects (stim and CTRL)
data.list = SplitObject(data, split.by = "ident")

# normalize and identify variable features for each dataset independently
data.list <- lapply(X = data.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = data.list)
```



```{r, eval = set_eval, echo=FALSE}
# Perform integration
# We then identify anchors using the FindIntegrationAnchors() function, which takes a list 
# of Seurat objects as input, and use these anchors to integrate the two datasets together with IntegrateData().
data.anchors <- FindIntegrationAnchors(object.list = data.list, anchor.features = features)

# this command creates an 'integrated' data assay
data.combined <- IntegrateData(anchorset = data.anchors)

# Save raw data
saveRDS(data.combined, file = paste("data_Integrated_5_samples.RDS") )
```



```{r, eval = set_eval, echo=FALSE}
# Load raw data
data.combined <- readRDS("data_Integrated_5_samples.RDS")
```


```{r, eval = set_eval, echo=FALSE}
# Run the standard workflow for visualization and clustering
data.combined <- ScaleData(data.combined, verbose = FALSE)
data.combined <- RunPCA(data.combined, npcs = 50, verbose = FALSE)
data.combined <- RunUMAP(data.combined, reduction = "pca", dims = 1:50)

# Save integrated data
saveRDS(data.combined, file = paste("data_Integrated_5_samples_umap.RDS") )
```


# `r colorize('UMAP for different samples before integration', 'blue')` {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Load non-integrated data
data.bf <- readRDS("data_raw_umap.RDS")

cat('##','All samples',' \n')
DimPlot(data.bf, reduction = "umap", group.by= 'sample' ) + ggtitle('All Samples')
cat(' \n \n')

for (samp in sample_Names_cp) {
  cat('##',samp,' \n')
  print(DimPlot(subset(data.bf, subset = (sample %in% c(samp))), reduction = "umap") + ggtitle(samp))
  cat(' \n \n')
}
```


```{r, echo=FALSE}
# Load integrated data
data.combined <- readRDS("data_Integrated_5_samples_umap.RDS")
```

# `r colorize('UMAP for different samples after integration', 'blue')` {.tabset}


```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('##','All samples',' \n')
DimPlot(data.combined, reduction = "umap", group.by= 'group' ) + ggtitle('All Samples')
cat(' \n \n')
for (sample in sample_Names_cp) {
  cat('##',sample,' \n')
  print(DimPlot(subset(data.combined, subset = (group %in% c(sample))), reduction = "umap", group.by= 'group' ) + ggtitle(sample))
  cat(' \n \n')
}
```

```{r, eval = set_eval, echo=FALSE}
# Plot and save integrated umap 
pdf(paste("Integrated_5_samples_umap.pdf"), width=10, height=8, useDingbats = FALSE)
DimPlot(data.combined, reduction = "umap", group.by = "group")
for (sample in sample_Names_cp) {
  print(DimPlot(subset(data.combined,subset = (group %in% c(sample))),reduction = "umap",group.by = "group"))
}
dev.off()
```
# `r colorize('Mk-primed gene set after integration (Rodriguez-Fraticelli et al., 2018a)', 'blue')` {.tabset}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library("xlsx")
# ###Fourteen_clst_markers = read.xlsx('shinobu_2023-03-02_trial_run_analysis.xlsx', sheetIndex = 1)
MK_markers = read.xlsx('shinobu_TableS2_GSA_gene_set-2.xlsx', sheetIndex = 'Table S2A')

library('stringr')

MK_markers_selected = str_to_title(MK_markers$A)

gene_names <- rownames(data.combined)

mask <- MK_markers_selected %in% gene_names
gene_selected = MK_markers_selected[mask]

```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
data.combined <- readRDS("data_Integrated_5_samples_umap.RDS")
DefaultAssay(data.combined) <- "RNA"
for (feature in gene_selected) {
  cat('##',feature,' \n')
  print(FeaturePlot(data.combined, features = c(feature),raster=FALSE, slot = "data"))
  print(VlnPlot(data.combined, features = c(feature), pt.size = 0.0,  assay = "RNA",
                group.by='nFeature_label')+ggtitle(paste(feature, '/ nFeature')))
  print(VlnPlot(data.combined, features = c(feature), pt.size = 0.0,  assay = "RNA", 
                group.by='nRead_label')+ggtitle(paste(feature, '/ nRead')))
  cat(' \n \n')
}
```
